ropenblas: Download, Compile and Link OpenBLAS Library with R

### 2021-03-19 paper/paper.Rmd
Update

# Statement of Need

The `ropenblas` package aims to allow algebraic computing of R to be performed using the OpenBLAS library, that is, it allows easy linking of R to the OpenBLAS library on GNU/Linux systems without depending on the distribution repositories. This will allow several researchers in the areas of statistics, data science, and machine learning to take advantage of more efficient performance in algebraic calculations, for example, multiplication, factorization, and matrix inversion. The `ropenblas` library, version 0.2.9 will also allow the R programmer to have, in his GNU/Linux distribution, several compiled versions of the R language giving the possibility to easily switch between these versions, this being an Open Source functionality that is only possible in some commercial IDEs of R. All this is done within the R language, minimizing the chance of less experienced users to break their operating system by running several instructions that are not perfectly understood.

The fact that the `ropenblas` package does not depend on the repositories of the GNU/Linux distribution will allow that in more stable distributions the R programmer will have at his disposal the most recent version of the OpenBLAS libraries and the R language. Everything is done safely, since the `ropenblas` package uses the stable versions of the official development repositories of the OpenBLAS library and the R programming language, respectively. Until the present version, the package has more than 11000 downloads, having an average of more than 1000 downloads in the month before the date of this submission, based on the official R language repositories.

@@

The term "computational efficiency" is very common for those who program statistical methods, in which a large part of them involve algebraic operations that are often reproduced in computationally intensive simulations, such as Monte-Carlo simulations - MC and resampling methods, as is the case with bootstrap resampling. Statistics is just one example within so many other areas that need performance and uses the R language.

In addition to the adoption of good programming practices and the maximum, efficient and adequate use of available computational resources, such as code parallelization, through multicore parallelism procedures allowed by most current processors and operating systems, small adjustments and linkage of libraries can provide useful benefits.

@@

The `ropenblas` package can be installed in two ways. The first is using the `install.packages()` function of the `utils` package which is available in any basic language installation and the second is using the `devtools` package which will allow the package to be installed directly from the development directory on GitHub.

@@

```{r bench, echo = FALSE, fig.cap = "Benchmarks of a decomposition of singular and inverse value of a matrix of dimension 1000 x 1000.", fig.align = "center", out.width="90%"}
knitr::include_graphics(path = "benchmark.png") 
```

>>>>

<!-- # Statement of Need -->

<!-- The `ropenblas` package aims to allow algebraic computing of R to be performed using the OpenBLAS library, that is, it allows easy linking of R to the OpenBLAS library on GNU/Linux systems without depending on the distribution repositories. This will allow several researchers in the areas of statistics, data science, and machine learning to take advantage of more efficient performance in algebraic calculations, for example, multiplication, factorization, and matrix inversion. The `ropenblas` library, version 0.2.9 will also allow the R programmer to have, in his GNU/Linux distribution, several compiled versions of the R language giving the possibility to easily switch between these versions, this being an Open Source functionality that is only possible in some commercial IDEs of R. All this is done within the R language, minimizing the chance of less experienced users to break their operating system by running several instructions that are not perfectly understood. -->

<!-- The fact that the `ropenblas` package does not depend on the repositories of the GNU/Linux distribution will allow that in more stable distributions the R programmer will have at his disposal the most recent version of the OpenBLAS libraries and the R language. Everything is done safely, since the `ropenblas` package uses the stable versions of the official development repositories of the OpenBLAS library and the R programming language, respectively. Until the present version, the package has more than 11000 downloads, having an average of more than 1000 downloads in the month before the date of this submission, based on the official R language repositories. -->

@@

<!-- The term "computational efficiency" is very common for those who program statistical methods, in which a large part of them involve algebraic operations that are often reproduced in computationally intensive simulations, such as Monte-Carlo simulations - MC and resampling methods, as is the case with bootstrap resampling. Statistics is just one example within so many other areas that need performance and uses the R language. -->

<!-- In addition to the adoption of good programming practices and the maximum, efficient and adequate use of available computational resources, such as code parallelization, through multicore parallelism procedures allowed by most current processors and operating systems, small adjustments and linkage of libraries can provide useful benefits. -->

@@

<!-- The `ropenblas` package can be installed in two ways. The first is using the `install.packages()` function of the `utils` package which is available in any basic language installation and the second is using the `devtools` package which will allow the package to be installed directly from the development directory on GitHub. -->

@@

```{r bench, echo = FALSE, fig.cap = "Benchmarks of a decomposition of singular and inverse value of a matrix of dimension 1000 x 1000.", fig.align = "center", out.width="60%"}
knitr::include_graphics(path = "benchmark.png") 
```
